Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra

### 2015-05-26 ms.tex
Modify Title and Abstract
From 3349de2:
============

In ms.tex,

(*) Modify Abstract
(*) Modify Title

\shorttitle{MWA Observations of Wide-field Effects in EoR Power Spectra}
\shortauthors{Thyagarajan et~al.}

@@

\title{Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra
using Murchison Widefield Array Observations}

@@

Here we confirm a recently predicted and previously unknown foreground signature in the 3D power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherently chromatic nature of a wide-field instrument response and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. Thus it is a critical input to design and analysis choices of future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array. The simulation in our recent paper which predicted this feature was validated against Murchison Widefield Array data but this key element was at or below the noise level. In this paper, we improve the Murchison Widefield Array data sensitivity through coherent averaging of 12 independent snapshots aligned in local sidereal time across different observing nights, and provide the first confirmation of the prediction with a signal-noise ratio $>10$.

>>>>

\shorttitle{Observations of Wide-field Effects in EoR Power Spectra}
\shortauthors{Thyagarajan et~al.}

@@

\title{Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra}

@@

We confirm the recently predicted ``pitchfork'' foreground signature in power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherent response of a wide-field instrument and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. In our recent paper, many signatures in the simulation which predicted this feature were validated against Murchison Widefield Array data but this key element was at or below the noise level. In this paper, we improve the data sensitivity through coherent averaging of 12 independent snapshots with identical instrument settings, and provide the first confirmation of the prediction with a signal-noise ratio $>10$. This wide-field effect can by mitigated by careful antenna design which suppresses sensitivity towards the horizon. With future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array, based on simple scaling relations, we estimate a suppression of $\gtrsim 40$~dB is required near the horizon for signal detection in cosmological H~{\sc i} modes in analyses that deploy the {\it foreground avoidance} technique.
